A Switching Planner for Combined Task and Observation Planning

### 2010-11-26 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
got new draft of switching, and Bayesian stuff for the evaluation section

Our approach always begins with a sequential session. Non-assumptive
actions from the trace are executed until the applicability of the
next scheduled action, here \action, is questionable. A contingent
session then begins, which tailors sensory processing by the agent to
perceive whether the assumptions made in the trace hold. For each
active assumption in the trace, we add a {\em disconfirm} action to
the POMDP whose execution is rewarding (resp. costly) if the
assumption is false (resp. true). We also add one {\em confirm} action
that is rewarding (resp. costly) to execute if $\poss(\action)$ is
true (resp. false). If execution of the contingent plan applies a
disconfirm action, then a new sequential session begins at the current
underlying belief-state. If the confirm action is executed, the
sequential session is resumed, and \action\ is executed. Otherwise,
control rests in the continual session.
%%
Finally, because contingent planning is only practical in relatively
small POMDPs, contingent sessions proceed in an abstract decision
process determined by the current trace. That abstract process is
constructed by first omitting all variables from the model that are
not featured in the trace, and by then iteratively refining that
abstraction while the resulting model is of a practicable size.

PLEASE STICK A PARAGRAPH HERE, (SENSIBLY) REQUESTED BY MORITZ, ABOUT
HOW WE DO BELIEF REVISION ACROSS SWITCHING.

@@

For sequential planning the operators from the POMDP model are
available with semantics that accommodate the above abstraction as
follows. In the deterministic model a proposition $\prop$ can be
thought to have a ternary interpretation at a state \state, as either
{\em true}, written $\prop\in\state$, {\em false}, written
$\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
notation written $\prop=\#\in\state$. For example, in $\pp{DORA}$
$(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
$\action\in\actions$ in the deterministic model, if
$\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
or negative effect of $\action$, then $\action$ is not applicable
in \state\ if $\prop=\#\in\state$. Indeed, first an assumptive action
must be applied to make a truth assignment to \prop, and then that
truth assignment could be modified by POMDP operators. Lastly, we
address the switching semantics of action execution. If
$\action\in\actions$ has a precondition that is not true with
probability $1$ in the underlying belief $\bstate$, then before it is
executed blindly, the system switches to a contingent session,
discussed in detail in the following section.

@@

In a decision-theoretic session, the switching planner solves a sensing
problem in an abstract processes defined in terms of the assumptive
actions in the current sequential plan, and the $(\pp{:init})$ declaration
from the original problem description.

The focus for decision-theoretic planning is on sensing, rather than
the achievement of utility according to the given objectives. We
therefor give the process in this session a different reward model, which
reflects the value of performing sensing actions in the context of the
trace proposed for the preceding sequential deliberation. In detail,
all positive rewards from the initial POMDP are removed. Then, for
each $\assumptiveS{i}$ action scheduled in the current sequential
plan, we have a {\em dual} $\assumptiveDT{i}$ so that
$\forall\state\in\states$:

@@

\noindent Where $\$(T_i)$ is a positive numeric quantity which
captures the reward the agent receives for correctly jettisoning a
false assumption. The added dual actions are propositionally trivial
-- i.e., $\forall\state\in\states$ $\poss(\assumptiveDT{i})
\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  Also, for the action,
$\action_i$, whose scheduled execution triggered the switch to the
decision-theoretic session, we include another propositionally trivial
action $\actions.\poss(\action_i)$ with the reward property:

@@

For the management of switching, execution of either a dual action of
the form $\assumptiveDT{i}$, or the $\actions.\poss(\action_i)$,
returns control to the sequential planner. At that point, if a dual is
executed, then the sequential planner must replan with $\bstate_0$
equal to the current belief. Otherwise, if $\actions.\poss(\action_i)$
is executed, then the current sequential plan is executed until
further sensing is scheduled, or to completion. In our project the
rewards, i.e., function $\$:2^\propositions\to\Re$, for the POMDP
posed to decision-theoretic planning are sourced from a motivational
component of our robotic architecture.

The starting belief-state for the abstract process is given in terms
of a {\em relaxed} $(\pp{:init})$ declaration.  We first construct a
tree which only includes terms from the {\em relaxed} visitation
characterised by the {\em assumptive} actions in the current
sequential plan. For example, if $\pp{DORA}$'s plan is:

\[
\begin{array}{l}
\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office});\\
(\pp{report-is-in}~\pp{box}~\pp{office})
\end{array}
\]

\noindent Then the {\em relaxed} declaration will be,

@@

\noindent At this point, decision-theoretic planning would proceed in an
abstraction of the environment that does not contain a $\pp{cup}$,
where the inclusion (exclusion) of POMDP actions is the same as we
defined it for the sequential session.

In a second step, we iteratively refine the above{\em relaxed}
declaration by adding terms from the original statement of
$(\pp{:init})$ while the resulting state space remains of a
practicable size. Of course, it is expensive to measure size exactly,
and in practice we estimate this to be $2^{|\propositions^\#|}$, where
$\propositions^\#$ is the set of propositions in the abstraction that
were not assigned the {\em unknown} value \#. Terms are proposed for
addition because they are probabilistic, and because they were omitted
in the {\em relaxed} visitation characterised by {\em assumptive}
actions in the sequential plan. In each iteration, we add one
candidate term with the largest number of parent terms already in the
candidate tree.

>>>>

Our approach always begins with a sequential session. Non-assumptive
actions from the trace are executed until the applicability of the
next scheduled action, here \action, is questionable. A contingent
session then begins, which tailors sensory processing by the agent to
perceive whether the assumptions made in the trace hold. For each {\em
active} assumption in the trace, we add a {\em disconfirm} action to
the POMDP whose execution is rewarding (resp. costly) if the
assumption is false (resp. true). We also add one {\em confirm} action
that is rewarding (resp. costly) to execute if $\poss(\action)$ is
true (resp. false). If execution of the contingent plan applies a
disconfirm action, then a new sequential session begins at the current
underlying belief-state. If the confirm action is executed, the
sequential session is resumed, and \action\ is executed. Otherwise,
control rests in the continual session.

Because contingent planning is only practical in relatively small
POMDPs, contingent sessions proceed in an abstract decision process
determined by the current trace. That abstract process is constructed
by first omitting all variables from the model that are not featured
in the trace, and by then iteratively refining that abstraction while
the resulting model is of a practicable size.

%% PLEASE STICK A PARAGRAPH HERE, (SENSIBLY) REQUESTED BY MORITZ, ABOUT
%% HOW WE DO BELIEF REVISION ACROSS SWITCHING.

@@

For sequential planning the operators from the POMDP model are
available with semantics that accommodate the above abstraction as
follows. In the deterministic model a proposition $\prop$ can be
thought to have a ternary interpretation at a state \state, as either
{\em true}, written $\prop\in\state$, {\em false}, written
$\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
$(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
$\action\in\actions$ in the deterministic model, if
$\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
or negative effect of $\action$, then $\action$ is not applicable
in \state\ if $\prop\#\in\state$. Indeed, first an assumptive action
must be applied to make a truth assignment to \prop, and then that
truth assignment could be modified by POMDP operators. Lastly, we
address the switching semantics of action execution. If
$\action\in\actions$ has a precondition that is not true with
probability $1$ in the underlying belief $\bstate$, then before it is
executed blindly, the system switches to a contingent session,
discussed in detail in the following section.

@@

In a contingent session our switching planner solves a sensing problem
in an abstract processes defined in terms of the assumptive actions in
the current trace, and the $(\pp{:init})$ declaration from the
original problem description.

The focus for contingent planning is on sensing, we therefore give the
process in this session an augmented reward model.  This reflects the
value of performing sensing actions in the context of the trace
proposed by the preceding sequential deliberation. First, all rewards
from the original POMDP are retained. Then, for each $\assumptiveS{i}$
action scheduled by the current trace, we have a {\em dual}
$\assumptiveDT{i}$ so that $\forall\state\in\states$:

@@

\noindent Where $\$(T_i)$ is a positive numeric quantity which
captures the reward the agent receives for correctly rejecting a
false assumption. The added dual actions are propositionally trivial
-- i.e., $\forall\state\in\states$ $\poss(\assumptiveDT{i})
\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  
%%
In order to reduce the number of possible rejections we do not
consider assumptions that are not {\em active} with respect to the
action, $\action_i$, whose scheduled execution switch the system into
the contingent session. A assumption $\assumptiveS{i}$ is active if
$\action_i$ is not applicable in the sequential model unless
$\assumptiveS{i}$ is scheduled in the trace prefix to $\action_i$. For
example, if $\pp{DORA}$'s trace is:

\[
\begin{array}{l}
\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}));\\
(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office});\\
(\pp{look-for-object}~\pp{DORA}~\pp{cup}~\pp{kitchen});\\
(\pp{report-is-in}~\pp{box}~\pp{office}); \\
(\pp{report-is-in}~\pp{cup}~\pp{kitchen})
\end{array}
\]

\noindent Taking the switching action $\action_i$ to be
$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$, we have that
$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
active, and therefore exclude
$\actions^{\bullet}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ from the
POMDP posed to the contingent session.

Again taking $\action_i$ to be the switching action, we include
another propositionally trivial action $\actions.\poss(\action_i)$
with the reward property:

@@

For the purposes of successive switching, execution of either a dual
action of the form $\assumptiveDT{i}$, or the
$\actions.\poss(\action_i)$, returns control to a sequential
session. If a dual is executed, then the sequential planner must
replan with $\bstate_0$ equal to the underlying
belief-state. Otherwise, if $\actions.\poss(\action_i)$ is executed,
then the current sequential plan is executed until further sensing is
scheduled, or to completion. In our project the rewards, i.e.,
function $\$:2^\propositions\to\Re$, for the POMDP posed to the
contingent session are sourced from a motivational component of our
robotic architecture.

The starting belief-state for the abstract process is given in terms
of a {\em relaxed} $(\pp{:init})$ declaration.  We first construct a
tree which only includes terms from the {\em relaxed} visitation
characterised by the active {\em assumptive} actions in the current
sequential plan. Taking our example trace from earlier, we have a {\em
relaxed} declaration:

@@

\noindent At this point, the session would proceed in an
abstraction of the environment that does not contain a
$\pp{cup}$. With regards to abstraction, the contingent and serial
sessions apply the same applicability conditions ---i.e., that each
effect and condition $\prop$ is specified $\prop\in\state$
(resp. $\prop\#\in\state$)--- to actions from the DTPDDL problem
description.

In a second step, we iteratively refine the relaxed declaration by
adding terms from the original statement of $(\pp{:init})$ while the
number of abstract states that occur with non-zero probability
according to that refined declaration remains of a practicable
size. In detail, for each excluded atomic term, we compute the {\em
entropy} of the corresponding proposition, {\em conditional} on the
active assumptions. The {\em conditional entropy} measure
is higher the more uncertain that proposition is given truth
assignments to propositions that are the subject of active
assumptions. Propositions that in the first place were excluded are
iteratively added to the problem posed to the conditional session in
increasing order according to that measure.

%% In particular, we compute the {\em conditional entropy} of
%% each excluded proposition given the active assumptions. Propositions
%% are candidates for addition in increasing order according to that
%% measure.

%% Of course, it is expensive to measure size exactly,
%% and in practise we estimate this to be $2^{|\propositions^\#|}$, where
%% $\propositions^\#$ is the set of propositions in the abstraction that
%% were not assigned the {\em unknown} value \#. Terms are proposed for
%% addition because they are probabilistic, and because they were omitted
%% in the {\em relaxed} visitation characterised by {\em assumptive}
%% actions in the sequential plan. In each iteration, we add one
%% candidate term with the largest number of parent terms already in the
%% candidate tree.
